---
title: "p8105_hw3_ph2538"
author: "Pei Yang Hsieh"
date: 2018-10-7
output: github_document
---

The purpose of this file is to complete homework 3.

```{r}
library(tidyverse)
library(ggplot2)
library(devtools)
library(p8105.datasets)
library(knitr)
library(patchwork)
library(ggridges)
```


## Problem 1

### Loading and cleaning brfss data 

```{r load_1}
#First, do some data cleaning:
#format the data to use appropriate variable names;
#focus on the “Overall Health” topic
#include only responses from “Excellent” to “Poor”
#organize responses as a factor taking levels from “Excellent” to “Poor”

data(brfss_smart2010)

clean_brfss =
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
  select(year, locationdesc, response, data_value) %>%
  mutate(response = as.factor(response)) %>%
  spread(key = response, value = data_value) %>%
  separate(locationdesc, into = c("state", "county"), sep = " - ")

```

### Question 1-4 for Problem 1

Using this dataset, do or answer the following (commenting on the results of each):

1. In 2002, which states were observed at 7 locations?

2. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

3. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

4. For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r state_7}
#In 2002, which states were observed at 7 locations?

brfss_2002 = filter(clean_brfss, year == '2002') %>%
  distinct(state, county) %>%
  count(state) %>%
  filter(n == 7)

brfss_2002
```

1. Connecticut, Florida, and North Carolina are observed at 7 locations.

```{r spaghetti_plot}
#Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

#count number of observations in each state
brfss_plot = clean_brfss %>%
  distinct(year, state, county) %>% 
  select(year,  state, county) %>% 
  group_by(year, state) %>%
  mutate(count = n())

#create plot
  ggplot(brfss_plot, aes(x = year, y = count, color = state)) +
  geom_point() + geom_line()
```

```{r excellent_table}
#Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

#the following takes mean and standard deviation by year, combining all NY counties in each year.
brfss_excellent = clean_brfss %>%
  filter(year == '2002' | year == '2006' | year == '2010') %>%
  filter(state == 'NY') %>%
  select(year, county, Excellent) %>%
  group_by(year) %>%
  summarize(sd_excellent = sd(Excellent), mean_excellent = mean(Excellent)) # %>%
  #spread(key = county, value = mean_excellent) %>%
  #knitr::kable(digits = 1)

brfss_excellent
```

```{r five_panel}
#For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

five_panel_plot = clean_brfss %>%
  mutate(year = as.factor(year)) %>% 
  group_by(year, state) %>%
  summarize(mean_excellent = mean(Excellent), mean_vgood = mean(`Very good`), mean_good = mean(Good), mean_fair = mean(Fair), mean_poor = mean(Poor)) %>%
  gather(key = response, value = proportion, mean_excellent:mean_poor) %>% 
  mutate(response_f = factor(response, levels = c('mean_excellent', 'mean_vgood', 'mean_good', 'mean_fair', 'mean_poor')))
  
ggplot(five_panel_plot, aes(x = year, y = proportion)) + 
  geom_boxplot() + 
  facet_grid(. ~ response_f) +
  theme(axis.text.x = element_text(angle = 90))
   
```

## Problem 2

```{r load_2}

#load instacart data
data(instacart)

instacart
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

The dimensions of the instacart dataset is`r dim(instacart)`, which means that it has 1384617 observations and 15 variables. Some key variables include order ID, product ID, and user ID which are all integers. There is also the product name, aisle, department, and department ID, which give information on the products and where they are found. 

### Questions for Problem 2

1. How many aisles are there, and which aisles are the most items ordered from?

2. Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

3. Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r aisle}
#How many aisles are there, and which aisles are the most items ordered from?

length(unique(instacart$aisle_id))

length(unique(instacart$aisle))

#there are 134 unique aisles, including an aisle called "missing"

aisle_max = 
  instacart %>%
  group_by(aisle_id) %>%
  summarize(n = n()) 

name_aisle_max = 
  instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) 

#sort by items sold
arrange(aisle_max, desc(n))

arrange(name_aisle_max, desc(n))

#Aisle 83 (fresh vegetables) has most items ordered from it, followed by aisle 24 (fresh fruits).

```

1. There are 134 unique aisles.

2. Aisle 83 (fresh vegetables) has most items ordered from it, followed by aisle 24 (fresh fruits).

```{r plot_aisle_number_items}
#Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
 
aisle_max %>% 
  mutate(aisle_id = as.factor(aisle_id)) %>% 
  mutate(aisle_id = forcats::fct_reorder(aisle_id, n)) %>% 
  ggplot(aes(x = aisle_id, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Items per Aisle",
    x = "Aisle ID",
    y = "Number of Items"
  ) +
  theme(axis.text.x = element_text(angle = 90), text = element_text(size = 5))


name_aisle_max %>% 
  mutate(aisle = as.factor(aisle)) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Items per Aisle",
    x = "Aisle Name",
    y = "Number of Items"
  ) +
  theme(axis.text.x = element_text(angle = 90), text = element_text(size = 5))

```

```{r popular_aisle}
#Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

popular_item = 
  instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  select(aisle, product_name) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
  
popular_item_sort = 
  popular_item %>%
  group_by(aisle) %>% 
  top_n(5) %>% 
  spread(key = aisle, value = product_name) %>% 
  knitr::kable(digits = 1)

popular_item_sort

```


```{r}
#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

apple_icecream = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = product_name, value = mean_hour) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Sunday", `1` = "Monday", `2` = "Tuesday", `3` = "Wednesday", `4` = "Thursday", `5` = "Friday", `6` = "Saturday")) %>%
   knitr::kable(digits = 1)
  
apple_icecream
```


## Problem 3

```{r load_3}
#load ny_noaa data
#Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

data(ny_noaa)

clean_noaa = 
  ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    prp = prcp/10,
    tmax = as.numeric(tmax)/10,
    tmin = as.numeric(tmin)/10
  )


```
### Questions for Problem 3

1. For snowfall, what are the most commonly observed values? Why?
2. Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
3. Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
#For snowfall, what are the most commonly observed values? Why?

clean_noaa %>% 
  group_by(snow) %>% 
  summarize(n = n()) %>% 
  top_n(5)

#0 is the most commonly observed value. This means that for the majority of the year, most days have no snowfall.
```
```{r}
#Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

plot_max_temp = 
  clean_noaa %>% 
  filter(month == "01" | month == "07") %>%
  na.omit() %>% 
  mutate(tmax = as.numeric(tmax)) %>% 
  group_by(year, month, id) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = mean_tmax)) + 
  geom_boxplot() + facet_grid(. ~ month) +
  labs(
    title = "Average Max Temp in January and July",
    x = "Year",
    y = "Average Max Temperature"
  ) +
  theme(axis.text.x = element_text(angle = 90))

plot_max_temp 

```
```{r}
#Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

tmax_tmin_plot = 
  clean_noaa %>% 
  na.omit() %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() +
  labs(
    title = "Minimum Temperature vs Max temperature",
    x = "Minimum Temperature",
    y = "Max Temperature"
  ) +
  theme(axis.text.x = element_text(angle = 90))
  
snowfall_distribution = 
  clean_noaa %>% 
  na.omit() %>% 
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = year, y = snow)) + 
  geom_boxplot() +
  labs(
    title = "Distribution of snowfall",
    x = "Year",
    y = "Snow"
  ) +
  theme(axis.text.x = element_text(angle = 90))

tmax_tmin_plot + snowfall_distribution

  
```

