---
title: "p8105_hw3_ph2538"
author: "Pei Yang Hsieh"
date: 2018-10-7
output: github_document
---

The purpose of this file is to complete homework 3.

```{r}
library(tidyverse)
library(ggplot2)
library(devtools)
library(p8105.datasets)
```


## Problem 1

### Loading and cleaning brfss data 

```{r load_1}
#First, do some data cleaning:
#format the data to use appropriate variable names;
#focus on the “Overall Health” topic
#include only responses from “Excellent” to “Poor”
#organize responses as a factor taking levels from “Excellent” to “Poor”

data(brfss_smart2010)

clean_brfss =
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
  select(year, locationdesc, response, data_value) %>%
  mutate(as.factor(response)) %>%
  separate(locationdesc, into = c("state", "county"), sep = " - ")

```

### Question 1-4 for Problem 1

Using this dataset, do or answer the following (commenting on the results of each):

1. In 2002, which states were observed at 7 locations?

2. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

3. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

4. For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r questions_1}

#In 2002, which states were observed at 7 locations?

brfss_2002 = filter(clean_brfss, year == '2002') %>%
  distinct(state, county) %>%
  count(state) %>%
  filter(n == 7)

brfss_2002
```

1. Connecticut, Florida, and North Carolina are observed at 7 locations.

```{r spaghetti_plot}

brfss_plot = clean_brfss %>%
  distinct(year, state, county) %>% 
  select(year,  state, county) %>% 
  group_by(year, state) %>%
  mutate(count = n())

#Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

  ggplot(brfss_plot, aes(x = year, y = count, color = state)) +
  geom_point() + geom_line()
```
```{r excellent_table}
#Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

brfss_excellent = clean_brfss %>%
  filter(year == '2002' | year == '2006' | year == '2010') %>%
  filter(state == 'NY') %>%
  select(year, state, county, response, data_value) %>%
  spread(key = response, value = data_value) %>%
  select(year, state, county, Excellent) %>%
  group_by(county, year) %>%
  summarize(sd_Excellent = sd(Excellent), mean_Excellent = mean(Excellent)) %>%
  spread(key = county, value = mean_Excellent) %>%
  knitr::kable(digits = 1)

```

```{r}
#For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

five_panel_plot = clean_brfss %>%
  group_by(year, county, state) %>%
  spread(key = response, value = data_value) %>%
  summarize(mean_Excellent = mean(Excellent!='NA'), mean_vgood = mean('Very good'!='NA'))
  

```

## Problem 2

```{r load_2}

#load instacart data
data(instacart)

```

## Problem 3

```{r load_3}

#load ny_noaa data
data(ny_noaa)




```

